Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-06-11 emsoft-2019/V2/response.tex

\Mahmoud{\section{Revision Letter}	
We thank the reviewers for the detailed and useful feedback on our submission. First, we will detail the main changes we made. Afterwards, we provide responses to your specific questions. 
\subsection{Changes}
We have improved our paper while taking into account all the minor comments. Particularly, \autoref{fig:overview} has been updated and the Remark \ref{rem:eps_Q} is added. Other changes are minor and are highlighted within the text.
\subsection{Responses}
Here, we provide answer to the main inquiries posed by reviewers. These points will be added later into the main body of the final version of our paper.\\
\textbf{Execution time for EMPCs (Reviewer 4)}\\
The total time to evaluate the PWA function for a given $x_k$ in \autoref{eq:affine_map} consists of two steps:
\begin{itemize}
\item Selecting the region $\mathcal R_i$ such that $x_k\in \mathcal R_i$.
\item To evaluate the corresponding affine map $F_i x_k +G_i$. 
\end{itemize}
Evaluating the affine map is straightforward and requires less computational effort compared to the region selection step. Direct implementation of region selection (\autoref{lst:caseof}) is simple, but inefficient. More efficiency can be achieved by leveraging binary search tree structures (e.g. \cite{Mnnigmann:2011}). In general, explicit MPC is well-known for its high-speed implementation over embedded systems with low computational power. \cite{Johansen:2007} shows that explicit MPCs with hundreds of regions can be implemented on application specific integrated circuit (ASIC) with about $20000$ gates, leading to computation times in the order of $1\mu s$. In \cite{Bemporad:2006}, it is shown that for typical problems evaluating the explicit MPC takes significantly shorter time in comparison to solving on-line QP. \\
\textbf{Termination of the algorithm for robust EMPC design (Reviewer 4)}\\
From theoretical point of view, the numerical error does not necessarily converge as each new designed controller may differ from the previous ones. Therefore, it is not easy to give an upper bound over the number of times that the robust MPC design needs to be repeated. However, one can use results on continuity of $argmin$ function in 
\begin{align*}
\label{eq:argmin}
&u_0(x_0,\Delta),\cdots,u_{N-1}(x_0,\Delta)=\\
&argmin \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
&\text{s.t.} \quad x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
%\quad &u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
&u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W}_{\Delta},\,\,\forall i\in\{0,1,\cdots,N\},
\end{align*}
 assuming its unique solution to every $\Delta$ and compute the Lipschitz constant for it over variations of $\Delta$. The computed Lipschitz constant will be only dependent on the dynamics of the system depicted in \autoref{eq:DSS} and hence can be used to evaluate the maximum error over incorrect region selection. This Lipschitz constant can be quite large, leading into a very conservative bound on maximum number of iterations before convergence. However, from experimental point of view, using a reasonable $\varepsilon_{SAFE}$, our observations show that convergence always happened after only a few iterations.
}

### 2019-06-11 emsoft-2019/V2/Cont_Syn.tex
updates

The third input parameter is $\varepsilon_Q$. In~\autoref{sec:Error_Analysis} we
showed that if the user provides a precision $p$, then we can automatically
compute $\varepsilon_Q$.
Now, we want to \emph{derive} a suitable precision $p$, however, and face an
issue. In order to derive $p$ for some expression, Daisy and any other precision
tuning tool requires an error bound which should be satisfied. I.e. either a
precision $p$ is given, and we can compute an error bound, or the error bound is
given and we can derive a precision $p$. We cannot do both at the same time; the
problem would be underconstrained. We solve this chicken-and-egg issue by
requiring the user to provide $\varepsilon_Q$ as the input. \Mahmoud{Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up-to $10^{-2}$. Increasing the $\varepsilon_Q$ leads into more memore save over $H$ and $K$, requiring more precision for $F$ and $G$ (\autoref{tab:di}). Of course, too large values for $\varepsilon_Q$ can result in failure of RMPC design. However, this is usually not the case over reasonable choices for $\varepsilon_Q$.}

>>>>

The third input parameter is $\varepsilon_Q$. In~\autoref{sec:Error_Analysis} we
showed that if the user provides a precision $p$, then we can automatically
compute $\varepsilon_Q$.
Now, we want to \emph{derive} a suitable precision $p$, however, and face an
issue. In order to derive $p$ for some expression, Daisy and any other precision
tuning tool requires an error bound which should be satisfied. I.e. either a
precision $p$ is given, and we can compute an error bound, or the error bound is
given and we can derive a precision $p$. We cannot do both at the same time; the
problem would be underconstrained. We solve this chicken-and-egg issue by
requiring the user to provide $\varepsilon_Q$ as the input. 
\Mahmoud{
	\begin{remark}\label{rem:eps_Q}
		Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up-to $10^{-2}$. Increasing the $\varepsilon_Q$ leads into more memory save over $H$ and $K$, while requiring more precision for $F$ and $G$ (\autoref{tab:di}). Of course, too large values for $\varepsilon_Q$ can result in failure of RMPC design. However, this is usually not the case over reasonable choices for $\varepsilon_Q$.
	\end{remark}}

### 2019-06-11 emsoft-2019/V2/Motiv.tex
updates

Next, we use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

>>>>

Next, we use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, \Mahmoud{for this introductory example,} the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

### 2019-06-11 emsoft-2019/V2/finite_precision.tex
updates

% mixed-precision tuning
To reduce memory usage and increase efficiency, we want to choose as short word
lengths as possible. Mixed-precision tuning tools automatically determine possibly different
word lengths for different values~\cite{DaisyTuning,Lee2006}. Compared to uniform precision (or word length),
mixed-precision often leads to improved resource usage, but due to the complexity of
fixed-point arithmetic and the error analysis, is challenging to do manually.  
Automated mixed-precision tuning tools usually perform a search: they repeatedly
select a candidate mixed-precision assignment (i.e. different word lengths for
different values) and check whether the assignment satisfies a given error
bound. The candidate assignments are chosen based on a heuristic which guides
the search towards promising candidates, e.g. using delta-debugging~\cite{DaisyTuning} or
simulated annealing~\cite{Lee2006}.

>>>>

% mixed-precision tuning
To reduce memory usage and increase efficiency, we want to choose as short word
lengths as possible, \Mahmoud{as this way memory footprint is minimized and run-time is not influenced.} Mixed-precision tuning tools automatically determine possibly different
word lengths for different values~\cite{DaisyTuning,Lee2006}. Compared to uniform precision (or word length),
mixed-precision often leads to improved resource usage, but due to the complexity of
fixed-point arithmetic and the error analysis, is challenging to do manually.  
Automated mixed-precision tuning tools usually perform a search: they repeatedly
select a candidate mixed-precision assignment (i.e. different word lengths for
different values) and check whether the assignment satisfies a given error
bound. The candidate assignments are chosen based on a heuristic which guides
the search towards promising candidates, e.g. using delta-debugging~\cite{DaisyTuning} or
simulated annealing~\cite{Lee2006}.
